\documentclass[letterpaper,11pt,twocolumn]{article}
\usepackage{geometry}
\usepackage{listings}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\title{CS574 Project Proposal}
\date{\today}
\hypersetup{
  pdfauthor={Kenny Ballou \& Marlin Roberts},
  pdftitle={Project Proposal},
  pdfkeywords={},
  pdfsubject={},
  pdflang={English}}
\author{Kenny Ballou \& Marlin Roberts}

\begin{document}

\maketitle{}
\tableofcontents{}

\section{Application}

We propose to examine the string constraint solving evaluation framework
located at:
\url{https://github.com/BoiseState/string-constraint-counting}~\cite{Sherman_2019}.
This codebase provides a framework for demonstrating the applicability of
various automata types used to represent symbolic strings in a string
constraint solving system.  It currently has implementations for the bounded,
acyclic and acyclic weighted automata types.

The major functionality requirement is for the framework to provide a
comparable set of outputs for the implemented automata.  This requires each
implementation of an automaton to perform operations in a comparable way that
matches the equivalent Java string operations.

A modified solver based on Java String Analyzer~(JSA)~\cite{strings2003} is
provided that can be used with any of the three automata types to solve various
benchmark constraints as well as a set of constraints obtained from analysis of
actual programs.  It can output satisfiability results as well as model count
results for comparing the accuracy of the automata models.

While the current codebase is largely compliant with the properties and
functionality identified, the instrumentation developed here could be used to
ensure that automata and automata factory classes developed in the future
adhere to the requirements of the framework.  With this in mind, we will try to
define properties on interfaces and abstract classes where possible.

\subsection{Metrics}

The approximate metrics of the application are shown in table
~\ref{tab:app-metrics}.


\begin{table}[ht]\label{tab:app-metrics}
  \centering
  \begin{tabular}{lr}
    \toprule
    Source Lines of Code & \(\sim14,000\) \\
    Source Files         & \(\sim70\) \\
    Classes              & \(\sim70\) \\
    Methods              & \(\sim700\) \\
    \bottomrule
  \end{tabular}
  \caption{Application Metrics}
\end{table}

\subsection{Areas of Interest}

\subsubsection{Interface Instrumentation vs. Implemented Class Instrumentation}

The bounded and acyclic automata types are implemented through an automata
modeling framework used by the Java String Analyzer~(JSA)~\cite{strings2003},
while the acyclic weighted implementation does not.  One of the aspects of
validation we can examine is the effectiveness of instrumenting an interface
versus a fully implemented class.  The surrounding \texttt{AutomatonModel}
class contains an instance of an automaton and is the implementation of an
abstract class, while the automaton itself does not implement an interface
other than \texttt{serializable} and \texttt{cloneable}.  We may be able to
introduce an interface for the automaton classes that does not affect operation
of the program but allows us have better control and placement of the
instrumentation.

\subsubsection{Modeling Behaviors}

Some of the desired behaviors involve the modeled String operations. Since the
framework is meant to be used to model Java String operations, the modeled
operations should perform as their Java counterpart does.  For instance, Java
String operation \texttt{substring~(start, end)} has specific behaviors that
depend on the values of \texttt{start}, \texttt{end}, and the length of the
string.  Specifically:

\begin{itemize}
\item{The \texttt{start} index is included in the returned string.}
\item{The \texttt{end} index is excluded from the returned string.}
\item{If \texttt{start} or \texttt{end} is negative, greater than the length of
    the string, or \texttt{start} is greater than \texttt{end}, then an
    exception is thrown.}
\end{itemize}

Any new type of automaton is expected to model these behaviors accurately in
order to be comparable to other types.  Since strings are represented by
automata, it may be difficult to express these requirements using JML and
assertions, since we cannot examine the returned values as an actual string.
Examining these automata may require introducing methods that are side-effect
free.  Existing methods used to test returned results will need to be examined
carefully for side-effects.

\section{Proposal of Changes}

We propose the usage of several verification tools to verify and otherwise
implement fixes for the following properties.

\subsection{Properties}

The following property classes have been identified to be areas of program
criticallity.  For each of the classes, we describe the nature of the set of
properties and the type of the property class.  For state based properties, we
will use either \texttt{Java Assertions}~\cite{oracle-java-assertions} or
\texttt{OpenJML}~\cite{openjml}.  For sequence based properties, we will use
\texttt{JavaMOP}~\cite{javamop}.

\paragraph{Automata Binary Operations} Operations involving two automata can
only be performed on automata of the same type.  We count roughly \(10\)
instances of this property.  Since this property is specified around operations
and not a sequence of operations, this is a state property.

\paragraph{Automata Return Types} Automata that return Automata need to
return the correct type.  There are about \(10\) instances of this property
that will need to be verified.  This property is state based since it is
centered around individual automata operations.

\paragraph{Automata Operations} Automata should be \textit{minimized} before
other operations occur.  Since this property is concerned with the ordering of
methods, this is a sequence property.

\paragraph{Automata Determinized Operations} Automata are required to be
\textit{determinized} before they are \textit{minimized}.  Automata operations
need to be in a proper order, therefore, this is a sequence property.

\paragraph{Automata Normalized Operations} Automata should be
\textit{normalized} before they are \textit{determinized}.  Similar to the
previous property class, the automata operations need to be verified via
sequential means.

\paragraph{String Operation Properties} Operations need to match Java String
operation behaviors, where it can be tested.  For example,
\texttt{substring~(start, end)} has a certain behavior associated with
(\texttt{start > end}) that should be replicated.  Specifically, the
\texttt{start} index is included in the returned string; the \texttt{end} index
is excluded from the returned string; and in the cases where \texttt{start} or
\texttt{end} are negative, \texttt{start} or \texttt{end} are greater than the
length of the string, or \texttt{start} is greater than \texttt{end} then
\texttt{substring} will raise an exception.  However, in the case of the
automata, an empty automaton is returned.  Similarly, \texttt{delete~(start,
  end)} has similar behaivor: the \texttt{start} index is included in the
removed string; \texttt{end} index is excluded from the removed string.
Furthermore, if the \texttt{end} index is greater than the length of the
string, all characters from the \texttt{start} index to the end of the string
are deleted.  If the \texttt{start} index is negative or greater than the
length of the string then \texttt{delete} raises an exception.  Again, however,
an empty automaton is returned instead of an exception.

There are about \(10\) unique instances of this property that will need to be
verified.  Operation behavior for certain argument values will have to be
determined and written into post-conditions of the operations.

\paragraph{Automata Types} Automata returned from operations should have certain
characteristics, if they can be tested in a side-effect-free way.  There are
about \(8\) unique instances of this property to be verified.

For example:

\begin{itemize}
\item{\texttt{concatenate}}
  \begin{itemize}
  \item{The maximum length of the string accepted by the returned automaton
      should be greater than or equal to the longest string of the original
      automata.}
  \end{itemize}
\item{\texttt{substring}}
  \begin{itemize}
  \item{The maximum string accepted by the returned automaton should be less
      than or equal to the lengths accepted by the original automata.}
  \end{itemize}
\end{itemize}

Other candidates to be considered:

\begin{itemize}
\item{\texttt{union}}
  \begin{itemize}
  \item{If either automata has final states, the union should have final
      states.}
  \item{The count of strings accepted by the automata should be greater or
      equal than the larger string count of the original automata.}
  \end{itemize}
\item{\texttt{intersection}}
  \begin{itemize}
  \item{The resulting automata may not contain final states.}
  \item{The count of strings accepted by the resulting automaton should be less
      than or equal to the lesser string count of the original automata.}
  \end{itemize}
\item{\texttt{insert}}
\item{\texttt{replace}}
  \begin{itemize}
  \item{The maximum length of the string accepted by the automaton should be
      unchanged.}
  \item{The count of strings accepted by the automaton should be unchanged.}
  \end{itemize}
\item{\texttt{delete}}
  \begin{itemize}
    \item{The maximum string length accepted by the resulting automaton should
        be less than or equal to the original automaton.}
  \end{itemize}
\item{\texttt{trim}}
\item{\texttt{reverse}}
  \begin{itemize}
  \item{The maximum string length accepted by the resulting automaton should be
      equal to that of the original automaton.}
  \item{The count of strings accepted by the resulting automaton should be unchanged.}
  \end{itemize}
\item{\texttt{assertions}}
  \begin{itemize}
    \item{Assertions work to reduce the possible set of strings accepted by the
        automaton.}
    \item{Resulting automata returned by \texttt{assertions} should have string
        lengths accepted be less than or equal to the strings accepted by the
        original automata.}
  \end{itemize}
\end{itemize}

\paragraph{Automata Trimming Operations} Automata are required to have dead states and epsilon
transitions removed before certain operations.

\paragraph{Solver/Timer Operations} Timers used by solvers should be started before stopped,
both before requesting the total run time.  Since this property is focusing on
the ordering of methods on timer objects, this is a sequential property.

\section{Conclusion}

The overall health of the String Constraint Counter framework will be better
off after the addition of these property verifications.  Furthermore, current
and future developers and researchers can proceed with more confidence as the
verification toolings can aid in ensuring new development maintains adherence
to the specifications.

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}
